Self-test prime dimensional EPR pairs with constant alphabet

### 2019-01-10 Draft/draft.tex
Honghao: the main proof should be more rigorous now.

\newcommand{\CHSH}{CHSH^{(d)}}
\newcommand{\MS}{MS}
\newcommand{\SVT}{SVT}
\newcommand{\EPR}[1]{EPR^{(#1)}}
\newcommand{\paulix}[1]{\sigma_x^{(#1)}}
\newcommand{\pauliz}[1]{\sigma_z^{(#1)}}
\newcommand{\G}[1]{G^{(#1)}}
\newcommand{\LS}{LS}
\newcommand{\tA}{\tilde{A}}
\newcommand{\tB}{\tilde{B}}
\newcommand{\tW}{\tilde{W}}
\newcommand{\tx}{\tilde{x}}
\newcommand{\tpsi}{\tilde{\psi}}
\newcommand{\tri}{\Delta}

@@

%-----------------------------------------------------------------
\subsection{Proof}
%-----------------------------------------------------------------
\hl{This subsection needs to be formalized.}
By Lemma~$4.3$ of \cite{coladan2017}, we can extract an operator solution from the perfect winning strategy 
of the linear system game: 
For each variable $\{ x_i \}_{i=1}^n$, Alice and Bob has operators $A_i$ and $B_i$ respectively.
The condition that they agree with assignment to variables means that 
\begin{align}
	\bra{\psi} A_i \otimes \overline{B_i} \ket{\psi} = 1 \Rightarrow A_i \otimes \overline{B_i} \ket{\psi} = \ket{\psi}
\end{align}
and the condition that Alice's assignments satisfy the constraint means that 
\begin{align}
	\Tr(\rho_A \Pi_{j: A_{ij} \neq 0} A_j) = \Tr(\rho_A) \text{ for all } 1 \leq i \leq m
\end{align}
where $\rho_A =  \Tr_B(\ketbra{\psi}{\psi})$, so for  any $\ket{v} \in \supp(\rho_A)$,
we have 
\begin{align}
\Pi_{j:A_{ij} \neq 0} A_j \ket{v} = \ket{v} \text{ for all } 1 \leq i \leq m.
\end{align}
Since the relation $uxu^\dagger = x^2$ is embedded in this linear system game, we know
\begin{align}
	A_3A_4 A_1A_2 (A_3A_4)^\dagger \ket{v}= (A_1A_2)^2 \ket{v} \text{ for all } \ket{v} \in \supp(\rho_A).
\end{align}
How $u$ and $x$ are splitted into order-$2$ generators can be found in Appendix~\ref{sec:construct}.
For simplicity, we define $X = A_1A_2$ and $U=A_3A_4$ such that
the condition is equivalent to
\begin{align}
	\label{eq:ux_relation}
	UXU^\dagger \ket{v} = X^2 \ket{v} \text{ for all } \ket{v} \in \supp(\rho_A).
\end{align}
We will come back to the implication of this condition later.

By \cref{lm:chsh_comp}, we can extract states $\ket{u_0}$ and $\ket{u_1}$ from the support of $\tA_\ast^\diamond$,   
and define $\ket{x_1} = 1/\sqrt{2}(\ket{u_0} + i\ket{u_1})$. 
\cref{prop:2d-subspace} implies that $\ket{x_1}$ is the eigenvector of $X$ with eigenvalue $e^{i\pi/d} = \omega_d$.
Similarly, define $\ket{x_{d-1}} = 1/\sqrt{2}(\ket{u_0} - i\ket{u_1})$ such that $X \ket{x_{d-1}} = \omega_d^{d-1}\ket{x_{d-1}}$.
Hence, by \cref{eq:ux_relation} we can calculate the eigen-decomposition of $X$ in $\supp(\rho_A)$.
as follows
\begin{align}
\label{eq:ladder}
 XU^\dagger \ket{x_1} = U^\dagger X^2 \ket{x_1} = \omega_d^2 U^\dagger \ket{x_1},
\end{align}
so $U^\dagger \ket{x_1}$ is an eigenvector of $X$ with eigenvalue $\omega_d^2$.
By induction, we know $X (U^\dagger)^i \ket{x_1} = \omega_d^{2^i} (U^\dagger)^i\ket{x_1}$. 
From the set $\{(U^\dagger)^i \ket{x_1}\}_{i=0}^{d-2}$, we can identify $\ket{x_i}$ for $i = 1 \dots  d-1$
such that $X \ket{x_i} = \omega_d^i \ket{x_i}$ and 
\begin{align}
S = \{ \ket{x_1}, \ket{x_2} \dots  \ket{x_{d-1}} \} \subset \supp(\rho_A).
\end{align}
On the other hand, we also know that $U \ket{x_i} = \ket{x_{i/2}}$ for $i = 1,2 \dots d-1$.
Note that $U$ and $X$ satisfy the condition of \cref{lm:ux_independ} when their actions are 
restricted to $\spn(S)$.

\hl{Assume $\supp(\rho_A) = \supp(\rho_B)$ and $\ket{\psi} = \vc(\sqrt{\rho})$.} 
The consistency condition is equivalent to
\begin{align}
	A_i \sqrt{\rho} B_i^\dagger = \sqrt{\rho}.
\end{align}
\hl{We further assume that $A_i = B_i$,} and define $\rho_S = \Pi_S \sqrt{rho} \Pi_S$
where $\Pi_S = \sum_{i=1}^{d-1} \ketbra{x_i}{x_i}$. 
Then we can conclude that 
\begin{align}
	U \rho_S U^{\dagger} = \rho_S && X \rho_S X^\dagger = \rho_S.
\end{align}
For any $k \in [d-1]$ and $l \in \{1,2\dots d-1\}$, the equation above tells us that 
\begin{align}
	U^kX^l \rho_S (U^kX^l)^\dagger = \rho_S.
\end{align}
By \cref{lm:ux_independ}, $\rho_S$ commutes with all the $(d-1) \times (d-1)$ matrices, which means that 
\begin{align}
	\rho_S \propto \Pi_S
\end{align}

In the end, we turn our attention back to the single value test. 
The implication of the single value test is that
\begin{align}
	A_\triangle^\diamond \ket{\psi} = A_\triangle^\diamond \x B_\triangle^\diamond \ket{\psi} = \frac{1}{\sqrt{d}}
	\ket{x_0}\ket{x_0}
\end{align}
where $X \ket{x_0} = \ket{x_0}$ and $\ket{x_0} \in \supp(\rho_A)$. 
Define $S^\perp = \{ \ket{x_0}\}$, then we know $\Pi_S^\perp \rho_A \Pi_S^\perp \propto \Pi_S$.
Hence we know $\rho_A$ acts as $\1$ on $S \cup S^\perp$, which yields that there exists i
sometries $V_A$ and $V_B$ such that
\begin{align}
	V_A \x V_B \ket{\psi} = \ket{\EPR{d}}.
\end{align}

>>>>

\newcommand{\CHSH}{CHSH^{(d)}}
\newcommand{\MS}{MS}
\newcommand{\SVT}{SVT}
\newcommand{\EPR}[1]{EPR^{(#1)}}
\newcommand{\paulix}[1]{\sigma_x^{(#1)}}
\newcommand{\pauliz}[1]{\sigma_z^{(#1)}}
\newcommand{\G}[1]{G^{(#1)}}
\newcommand{\LS}{LS}
\newcommand{\tA}{\tilde{A}}
\newcommand{\tB}{\tilde{B}}
\newcommand{\tW}{\tilde{W}}
\newcommand{\tx}{\tilde{x}}
\newcommand{\tpsi}{\tilde{\psi}}
\newcommand{\tri}{\Delta}
\newcommand{\lB}{\overline{B}}

@@

%-----------------------------------------------------------------
\subsection{Proof}
%-----------------------------------------------------------------
\hl{This subsection needs to be formalized.}
By Lemma~$4.3$ of \cite{coladan2017}, we can extract an operator solution from the perfect winning strategy 
of the linear system game: 
For each variable $\{ x_i \}_{i=1}^n$, Alice and Bob has operators $A_i$ and $B_i$ respectively.
The condition that they agree with assignment to variables means that 
\begin{align}
	\bra{\psi} A_i \otimes \overline{B_i} \ket{\psi} = 1 \Rightarrow A_i \otimes \overline{B_i} \ket{\psi} = \ket{\psi}
\end{align}
and the condition that Alice's assignments satisfy the constraint means that 
\begin{align}
	\Tr(\rho_A \Pi_{j: A_{ij} \neq 0} A_j) = \Tr(\rho_A) \text{ for all } 1 \leq i \leq m
\end{align}
where $\rho_A =  \Tr_B(\ketbra{\psi}{\psi})$. 
Similarly we define $\rho_B = \Tr_A(\ketbra{\psi}{\psi})$.
For any $\ket{v} \in \supp(\rho_A)$,
we have 
\begin{align}
\Pi_{j:A_{ij} \neq 0} A_j \ket{v} = \ket{v} \text{ for all } 1 \leq i \leq m.
\end{align}
Since the relation $uxu^\dagger = x^2$ is embedded in this linear system game, we know
\begin{align}
	A_3A_4 A_1A_2 (A_3A_4)^\dagger \ket{v}= (A_1A_2)^2 \ket{v} \text{ for all } \ket{v} \in \supp(\rho_A).
\end{align}
How $u$ and $x$ are splitted into order-$2$ generators can be found in Appendix~\ref{sec:construct}.
For simplicity, we define $X_A = A_1A_2$ and $U_A=A_3A_4$ such that
the condition is equivalent to
\begin{align}
	\label{eq:ux_relation}
	U_AX_AU_A^\dagger \ket{v} = X_A^2 \ket{v} \text{ for all } \ket{v} \in \supp(\rho_A).
\end{align}
We will come back to the implication of this condition later.

By \cref{lm:chsh_comp}, we can extract states $\ket{u_0}$ and $\ket{u_1}$ from the support of $A_\ast^\diamond$,   
and define $\ket{x_1} = 1/\sqrt{2}(\ket{u_0} + i\ket{u_1})$. 
\cref{prop:2d-subspace} implies that $\ket{x_{A,1}}$ is the eigenvector of $X_A$ with eigenvalue $\omega_d$.
Similarly, define $\ket{x_{A,d-1}} = 1/\sqrt{2}(\ket{u_0} - i\ket{u_1})$ such that $X_A \ket{x_{A,d-1}} = \omega_d^{d-1}\ket{x_{A,d-1}}$.
Hence, by \cref{eq:ux_relation} we can calculate the eigen-decomposition of $X_A$ in $\supp(\rho_A)$.
as follows
\begin{align}
\label{eq:ladder}
 X_AU_A^\dagger \ket{x_{A,1}} = U_A^\dagger X_A^2 \ket{x_{A,1}} = \omega_d^2 U_A^\dagger \ket{x_{A,1}},
\end{align}
so $U_A^\dagger \ket{x_{A,1}}$ is an eigenvector of $X_A$ with eigenvalue $\omega_d^2$.
By induction, we know $X_A (U_A^\dagger)^i \ket{x_{A,1}} = \omega_d^{2^i} (U_A^\dagger)^i\ket{x_{A,1}}$. 
From the set $\{(U_A^\dagger)^i \ket{x_{A,1}}\}_{i=0}^{d-2}$, we can identify $\ket{x_{A,i}}$ for $i = 1 \dots  d-1$
such that $X_A \ket{x_{A,i}} = \omega_d^i \ket{x_{A,i}}$ and 
\begin{align}
S_A = \spn \{ \ket{x_{A,1}}, \ket{x_{A,2}} \dots  \ket{x_{A,d-1}} \} \subset \supp(\rho_A).
\end{align}
On the other hand, we also know that $U_A \ket{x_{A,i}} = \ket{x_{A,i/2}}$ for $i = 1,2 \dots d-1$.
Note that $U_A$ and $X_A$ satisfy the condition of \cref{lm:ux_independ} when their actions are 
restricted to $\spn(S_A)$.

Similar argument applies to Bob's operators. We can identify $X_B = B_1B_2$, 
$U_B=B_3B_4$ and a vector space $S_B =\spn\{ \ket{x_{B,1}}, \ket{x_{B,2}} \dots  \ket{x_{B,d-1}} \} \subset \supp(\rho_B)$,
such that 
\begin{align}
	X_B = \sum_{i=1}^{d-1} \omega_d^i \ketbra{x_{B,i}}{x_{B,i}} && U_B = \sum_{i=1}^{d-1} \ketbra{x_{B,i/2}}{x_{B,i}},
\end{align}
when their actions are restricted to $S_B$. Moreover, we can define a unitary $V: \supp{\rho_B} \to \supp(\rho_A)$,
which maps $\ket{x_{B,i}}$ to $\ket{x_{A,i}}$.

%and projectors $\Pi_{S_A}$ and $\Pi_{S_B}$ to be the projectors onto $S_A$ and $S_B$ respectively.
%From now on, we focus on the state $\ket{\psi_S}$ which is defined as
%\begin{align}
%	\ket{\psi_S} = (\Pi_{S_A} \x (V\Pi_{S_B}))\ket{\psi} = \vc(\tau) 
%\end{align}
%for some $\tau \in L(S_A)$.
Assume that $\ket{\psi} = \vc(\tau)$ for some $\tau \in L(\supp(\rho_B) \to \supp(\rho_A))$
The consistency condition is equivalent to
\begin{align}
\label{eq:con_tau}
	A_i \tau B_i^\dagger = \tau.
\end{align}
Substituting $i=1,2$ into \cref{eq:con_tau}, we get 
\begin{align}
	X_A \tau X_B^\dagger = A_1A_2\tau B_2^\dagger B_1^\dagger = A_1\tau B_1^\dagger = \tau.
\end{align}
Inserting $V^\dagger V$ between $\tau$ and $X_B$ and multiplying $V^\dagger$ to the right on both sides yield that 
\begin{align}
	X_A \tau V^\dagger X_A^\dagger = X_A \tau V^\dagger VX_B^\dagger V^\dagger = \tau V^\dagger,
\end{align}
where we use the relation $VX_B^\dagger V^\dagger = X_A$.
Similar argument gives us that 
\begin{align}
	U_A \tau V^\dagger  U_A^\dagger = \tau V^\dagger.
\end{align}
Then we can conclude that for any $k \in \{0,1 \dots d-2\}$ and $l \in \{1,2\dots d-1\}$
\begin{align}
	U_A^kX_A^l \tau V^\dagger (U_A^kX_A^l)^\dagger = \tau V^\dagger.
\end{align}
By \cref{lm:ux_independ}, $\tau V^\dagger$ projected onto $L(S_A)$ commutes with all the $(d-1) \times (d-1)$ matrices, which means that 
\begin{align}
	\label{eq:d-1}
	\tau V^\dagger|_{S_A}  \propto \1_{S_A}.
\end{align}

In the end, we turn our attention back to the single value test. 
The implication of the single value test is that
\begin{align}
	A_\triangle^\diamond \ket{\psi} = A_\triangle^\diamond \x B_\triangle^\diamond \ket{\psi} = \frac{1}{\sqrt{d}}
	\ket{x_{A,0}}\ket{x_{B,0}}
\end{align}
where $X_C \ket{x_{C,0}} = \ket{x_{C,0}}$ and $\ket{x_{C,0}} \in \supp(\rho_C)$ for $C = A$ or $B$. 
Recalling \cref{eq:d-1}, we can conclude that there exists local isometries 
$V_A$ and $V_B$ such that
\begin{align}
	V_A \x V_B \ket{\psi} = \ket{\EPR{d}}.
\end{align}
